Quantifying stochastic noise in cultured\\ circadian reporter cells

### 2015-03-13 main_text.tex
removing figs
wrote cover, need to do author summary, move methods after results
adding author summary

\pagebreak
\begin{abstract}
Stochastic noise at the cellular level has been shown to play a fundamental role in circadian oscillations, influencing how groups of cells entrain to external cues and likely serving as the mechanism by which cell-autonomous rhythms are generated.
Despite this importance, few studies have investigated how clock perturbations affect stochastic noise - even as increasing numbers of high-throughput screens categorize how gene knockdowns or small molecules can change clock period and amplitude.
This absence is likely due to the difficulty associated with measuring cell-autonomous stochastic noise directly, which currently requires the careful collection and processing of single-cell data.
% since cultured circadian reporters lack significant intercellular coupling,
In this study, we show that the damping rate of population-level bioluminescence recordings can serve as an accurate measure of cell-autonomous stochastic noise, and one which can be applied to future and existing high-throughput circadian screens.
Using cell-autonomous fibroblast data, we first show directly that higher noise at the single-cell results in faster damping at the population level.
Next, we show that the damping rate of cultured cells can be changed in a dose-dependent fashion by small molecule modulators, and confirm that such a change can be explained by single-cell noise using a mathematical model.
We further demonstrate the insights that can be gained by applying our method to a genome-wide siRNA screen, revealing that stochastic noise is altered independently from period, amplitude, and phase.
Finally, we hypothesize that the unperturbed clock is highly optimized for robust rhythms, as very few gene perturbations are capable of simultaneously increasing amplitude and lowering stochastic noise.
Ultimately, this study demonstrates the importance of considering the effect of circadian perturbations on cell-autonomous noise, particularly with regard to the development of small-molecule circadian therapeutics.
\end{abstract}

>>>>

\pagebreak
\begin{abstract}
Stochastic noise at the cellular level has been shown to play a fundamental role in circadian oscillations, influencing how groups of cells entrain to external cues and likely serving as the mechanism by which cell-autonomous rhythms are generated.
Despite this importance, few studies have investigated how clock perturbations affect stochastic noise - even as increasing numbers of high-throughput screens categorize how gene knockdowns or small molecules can change clock period and amplitude.
This absence is likely due to the difficulty associated with measuring cell-autonomous stochastic noise directly, which currently requires the careful collection and processing of single-cell data.
In this study, we show that the damping rate of population-level bioluminescence recordings can serve as an accurate measure of cell-autonomous stochastic noise, and one which can be applied to future and existing high-throughput circadian screens.
Using cell-autonomous fibroblast data, we first show directly that higher noise at the single-cell results in faster damping at the population level.
Next, we show that the damping rate of cultured cells can be changed in a dose-dependent fashion by small molecule modulators, and confirm that such a change can be explained by single-cell noise using a mathematical model.
We further demonstrate the insights that can be gained by applying our method to a genome-wide siRNA screen, revealing that stochastic noise is altered independently from period, amplitude, and phase.
Finally, we hypothesize that the unperturbed clock is highly optimized for robust rhythms, as very few gene perturbations are capable of simultaneously increasing amplitude and lowering stochastic noise.
Ultimately, this study demonstrates the importance of considering the effect of circadian perturbations on cell-autonomous noise, particularly with regard to the development of small-molecule circadian therapeutics.
\end{abstract}

\section*{Author Summary}

As most organisms exist in an environment which changes predictably with a 24-hour period, highly optimized genetic circuits turn on and off the production of key regulatory proteins to anticipate the day/night cycle.
In humans, the demands of a modern society have required that we deviate from this evolutionarily prescribed sleep and feeding schedule, resulting in increased long-term risks of metabolic disease.
There is therefore a desire to find pharmacological treatments which would restore the normal functioning of our circadian clock despite irregular behavioral schedules.
One aspect of these treatments which is often overlooked in searching for candidate drugs is how these treatments might affect the accuracy of the circadian timing system.
% Since each cell in the body keeps its own internal time, increasing the sloppiness of the clock machinery can result in tissues not being able to agree on what time to turn on protein production.
Recording the time of each cell is possible but difficult; as a result single-cell approaches cannot be scaled up to high-throughput searches.
In this paper, we show that it is possible to estimate how much the noise of a system has changed by looking only at the averaged protein production of an entire population of cells.
Such an approach allows us to analyze prior data for high-throughput screens, and show that the unperturbed clock has been highly optimized to be both accurate and high amplitude.
